Formalizing Stack Safety as a Security Property

### 2023-02-02 submission5/intro-v2.tex
Quick pass on I+II

Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19} not to categorically
rule out attacks. Other mechanisms are stronger, but relatively few formally state their guarantees.
The Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer (various) stack safety formalizations.

@@

In particular, many enforcement techniques focus purely on
well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
to enforce it completely. Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
to the combination of \(\clri\) and \(\clrc\). It is not until Georges et al. \cite{Georges22:TempsDesCerises}
that the Cerise line of work considers confidentiality; they also consider \(\clec\), but not
\(\clei\). \sna{It is really tempting to find a way to take credit for this.}

@@

Our artifact contains Coq formalizations of our properties for illustrative purposes
as well as their testing implementation. There are no proofs---we use primarily Coq for the
QuickChick library and to ensure that the formalizations are unambiguous.

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers are treated appropriately by whichever compiler
generated their code. Our properties explicitly require them to be treated as belonging
to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.

@@

This is a strong threat model, but it does omit some important aspects of
security enforcement: in particular, hardware and timing attacks are out of scope, and
we do not address concurrency.

>>>>

Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically
rule out attacks. Other mechanisms are stronger, but relatively few formally state their guarantees.
The Cerise work of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer (various) stack safety formalizations.

@@

In particular, many enforcement techniques focus purely on
well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
to enforce it completely. Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
to the combination of \(\clri\) and \(\clrc\). It is not until Georges et al. \cite{Georges22:TempsDesCerises}
that the Cerise line of work considers confidentiality; they also consider \(\clec\), but not
\(\clei\). \sna{It is really tempting to find a way to take credit for this.}
%
\rb{This is the direct support I can find: ``Following Anderson et al. [2021],
we must distinguish between {\em local state integrity} and {\em local state
confidentiality}.'' + similar discussion in their related work + their novel
claim of adding support for confidentiality properties, thus ``following'' us
\ldots + personal communication?}

@@

Our artifact contains Coq formalizations of our properties for illustrative purposes
as well as their testing implementation. There are no proofs---we use primarily Coq for the
QuickChick library and to ensure that the formalizations are unambiguous.
%
\rb{Mention proofs in future work?}

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers are treated appropriately by whichever compiler
generated their code. Our properties explicitly require them to be treated as belonging
to the caller, which could be enforced by a micro-policy or by a well-behaved compiler.
%
\rb{This last bit about callee-saved registers belonging to the caller is subtle, I think it risks the reader missing the point.}

@@

This is a strong threat model, but it does omit some important aspects of
security enforcement: in particular, hardware and timing attacks are out of scope, and
we do not address concurrency.
%
\rb{Anything about concurrency in future work, though?}
